\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2019-04-08 archive.tex
discussion progress

Our model is structured
around the hypothesis that
domain expertise
consists primarily of two ingredients.
First, domain experts have 
explicit, declarative concepts that are powerful yet finely-tuned.
A visual artist has concepts like arcs, spirals, symmetries, and perspectives;
a physicist has concepts like dot products, vector fields, and conservation laws;
and an architect has concepts like arches, supports, and bridges.
Second, experts have implicit procedural skill in deploying those concepts
to quickly solve new problems:
at a glance, human domain
experts can intuit which compositions of concepts are likely to solve
the task at hand, even before searching for a solution.
Studies of human expertise
find that,
compared to novices,
experts can quickly categorize and classify problems based on the ``deep structure''
of the problem's solution~\cite{chi1981expertise,chi2012seeing}.
In short, we take the stance that expertise means both
having the right concepts, and being able to quickly see how those concepts can be composed into a solution.

Because solutions to many kinds of
problems can often be described as some kind of
program~\cite{lake2015human,Ullman2012,DBLP:journals/cogsr/SchmidK11},
our model solves a specific problem by synthesizing a program solving it. It
gradually grows its domain-specific knowledge for a class of problems
by assembling a library of code containing concepts useful for the
domain, and by training a neural network to quickly infer, for a
specific problem, what code in its library is likely to
solve it. We think of these two learned components as analogous
to the explicit concepts and
implicit procedural skill that human experts develop.
We will first investigate our model within classic program
synthesis domains for manipulating sequences of numbers and text, and
then consider visual and creative programs for drawing pictures and
building towers out of toy blocks, and finally consider programs for basic kinds of equation discovery and programming language design, altogether considering both
deterministic and probabilistic programs that act both generatively
(e.g., producing an artifact like an image or plan) and conditionally
(e.g., mapping inputs to outputs).

@@

@@

\noindent\textbf{\system Hyperparameters}
Due to the high computational cost we
performed only an informal coarse
hyperparameter search.
The most important parameter is
the enumeration timeout during the wake phase;
domains that present more challenging program synthesis
problems require either longer timeouts, more CPUs, or both.
\begin{center}
  \begin{tabular}{ccccccc}
    \toprule
    Domain&Timeout&CPUs&Batch size&$\lambda$ (\ref{appendixCompression})&$\alpha$ (\ref{mapAppendix})&Max frontier (\ref{systemPseudocode})\\\midrule
    Lists&7m&64&10&1&30&5\\
    Text&7m&64&10&1&30&5\\
    Graphics&1h&128&40&1.5&30&5\\
    Symbolic regression&2m & 40&10 & 1&30&5\\
    Towers& 2m& 64&10 & 1&30&5\\
    Regexes&30m & 64&40 & 1&30&5\\
    \bottomrule     
  \end{tabular}
\end{center}

>>>>

Our model is structured
around the hypothesis that
domain expertise
consists primarily of two ingredients.
First, domain experts have 
explicit, declarative concepts that are powerful yet finely-tuned.
A visual artist has concepts like arcs, spirals, symmetries, and perspectives;
a physicist has concepts like dot products, vector fields, and conservation laws;
and an architect has concepts like arches, supports, and bridges.
Second, experts have implicit, procedural skill in deploying those concepts
to quickly solve new problems:
at a glance, human domain
experts can intuit which compositions of concepts are likely to solve
the task at hand, even before searching for a solution.
Studies of human expertise
find that,
compared to novices,
experts can quickly categorize and classify problems based on the ``deep structure''
of the problem's solution~\cite{chi1981expertise,chi2012seeing}.
In short, we take the stance that expertise means both
having the right explicit concepts, and being able to quickly see how those concepts can be composed into a solution.

Because solutions to many kinds of problems can often be described as
some kind of
program~\cite{lake2015human,Ullman2012,DBLP:journals/cogsr/SchmidK11},
our model solves a specific problem by synthesizing a program solving
it. It gradually grows its explicit domain-specific knowledge for a
class of problems by assembling a library of code containing concepts
useful for the domain, and by training a neural network to quickly
infer, for a specific problem, what code in its library is likely to
solve it. We think of these two learned components as analogous to the
explicit concepts and implicit procedural skill that human experts
develop.  The model is an instance of what in the machine learning
community is called a wake-sleep algorithm, where for us, `waking'
corresponds to solving problems, while a pair of `sleep' cycles
correspond to improving the library of explicit concepts and training
the neural net to improve implicit procedural knowledge of how to
write code.

We will first investigate our model within classic program
synthesis domains for manipulating sequences of numbers and text, and
then consider visual and creative programs for drawing pictures and
building towers out of toy blocks, and finally consider programs for basic kinds of equation discovery and programming language design, altogether considering both
deterministic and probabilistic programs that act both generatively
(e.g., producing an artifact like an image or plan) and conditionally
(e.g., mapping inputs to outputs).

@@

\emph{(what we found)} Our goal was to build a model capturing
humanlike aspects of learning to be an expert in a new domain of
problems.  Two main findings of our work are that, first, a single
program-induction system can learn to solve large sets of problems
from many qualitatively different domains; and second, that fully
acquiring expertise in these domains hinges both upon explicit
declarative knowledge, and implicit procedural skill.  We have
deliberately placed the interplay of declarative and procedural
knowledge at the center of our model, and believe that both these
kinds of learning are crucial for building agents that, like humans,
autonomously learn to navigate a new domain of problems.

Our model's domain knowledge grows over a series of wake/sleep cycles,
with the solutions to easier problems bootstrapping
solving of harder tasks.
This dynamic is related to, but distinct from, curriculum learning. In curriculum learning approaches an agent solves a stream of pedagogically selected, increasingly difficult tasks; here we consider the case with the agent lacks a `teacher' and instead sees randomly sampled batches of tasks, effectively self-pacing its way through an implicit curriculum.
But humans select their tasks in much richer ways,
and can even generate their own tasks to solve,
either as stepping stones toward hard, unsolved problems or motivated by curiosity and aesthetics.
Building agents that generate their own problems in these humanlike ways is a necessary next step
if we want machines that push against boundary of human domain knowledge or which even discover whole new domains of problems,
both of which go far beyond the relatively small-scale demonstrations in this work.

A basic representational challenge is to
learn concepts fundamentally inexpressible in terms of the starting basis: \system invents procedures for
adding vectors, taking dot products, and calculating forces,
but can not make the jump needed to define new terms like `acceleration' and `coordinate system,'
concepts inexpressible as a library of procedures.
We believe approaches based either on term rewriting systems or defining new generalized algebraic datatypes could give traction to fundamental problems like these,
strategies which are closely allied with conceptual role semantics.
Indeed, humans often open up whole new fields of inquiry by introducing concepts not
definable in terms of prior abstractions,
like the meaning of `allele' in genetics
or `force' in physics.

related work

limitations

how will go beyond? generate own tasks; real DSL learning (all expressible in base language); lifelong learning (forgetting?)

@@

\noindent\textbf{\system Hyperparameters}
Due to the high computational cost we
performed only an informal coarse
hyperparameter search.
The most important parameter is
the enumeration timeout during the wake phase;
domains that present more challenging program synthesis
problems require either longer timeouts, more CPUs, or both.
\begin{center}
  \begin{tabular}{ccccccc}
    \toprule
    Domain&Timeout&CPUs&Batch size&$\lambda$ (\ref{appendixCompression})&$\alpha$ (\ref{mapAppendix})&Max beam size (\ref{systemPseudocode})\\\midrule
    Lists&7m&64&10&1.5&30&5\\
    Text&7m&64&10&1.5&30&5\\
    Graphics&1h&96&50&1.5&30&5\\
    Symbolic regression&2m & 40&10 & 1&30&5\\
    Towers& 1h& 64&50 & 1.5&30&5\\
    Regexes&30m & 64&40 & 1.5&30&5\\
    \bottomrule     
  \end{tabular}
\end{center}
